---
title: "Project 1"
output:
    github_document:
      toc: true
    html_document:
      toc: true
---

## Link to repo

[GIT repo](https://github.com/hkopanski/ST_558_Proj_1)

## Set up of API Script
The following libraries are used in this script

```{r setup, message = FALSE, error= FALSE}
library(httr)
library(jsonlite)
library(tidyverse)
library(DT)
library(RSQLite)
library(DBI)
library(knitr)
```

## Create data functions

The following functions pull data from two sources:

https://statsapi.web.nhl.com/api/v1

and 

https://records.nhl.com/site/api

The data can be parsed during the request using modifiers or it can be parsed afterwards. There are 6 functions in this section:

- get_db_stats: to pull teams data from statsapi.web.nhl.com
- get_db_records: to pull franchise data from records.nhl.com
- get_records: to pull franchise details data from records.nhl.com
- get_goalie_data: pull goalie stats for a particular team
- get_skater_data: pull skater stats for a particular team
- get_team_stats2: pull team stats for a specific team

All functions output the data in the form of a tibble.

```{r functions, message = FALSE, error= FALSE}
get_db_stats <- function(endpoint = NULL, modifier = NULL){
  base <- "https://statsapi.web.nhl.com/api/v1"
  
  URL <- paste0(base, "/" , endpoint, modifier)
  temp_con <- GET(URL)
  temp_text <- content(temp_con, "text", encoding = "UTF-8")
  temp_JSON <- fromJSON(temp_text, flatten = TRUE)
  return(as_tibble(temp_JSON))
}

get_db_records <- function(endpoint = NULL, modifier = NULL){
  base <- "https://records.nhl.com/site/api"
  
  URL <- paste0(base, "/" , endpoint, modifier)
  temp_con <- GET(URL)
  temp_text <- content(temp_con, "text", encoding = "UTF-8")
  temp_JSON <- fromJSON(temp_text, flatten = TRUE)
  return(as_tibble(temp_JSON))
}

# This is to pull team specific information
get_records <- function(team = NULL, ID = NULL){
  df_franchise_det <- get_db_records('franchise-detail')
  df_franchise_det <- df_franchise_det$data %>% rowwise() %>%
    mutate(short_name = tail(strsplit(teamFullName, " ")[[1]], n = 1))
    if(is.null(team) & !is.null(ID)){
        temp_string <- paste0('franchise-season-records?cayenneExp=franchiseId=', as.character(ID))
        get_db_records(temp_string)}
    else if(!is.null(team)){
        team_alt = tail(strsplit(team, " ")[[1]], n = 1)
        temp_ID <- df_franchise_det %>% filter(toupper(short_name) == toupper(team_alt)) %>% select(id)
        temp_string <- paste0('franchise-season-records?cayenneExp=franchiseId=', as.character(temp_ID))
        get_db_records(temp_string)
    }
}

# This pulls goalie information
get_goalie_data <- function(team = NULL, ID = NULL){
  df_franchise_det <- get_db_records('franchise-detail')
  df_franchise_det <- df_franchise_det$data %>% rowwise() %>%
    mutate(short_name = tail(strsplit(teamFullName, " ")[[1]], n = 1))
  if(is.null(team) & !is.null(ID)){
    temp_string <- paste0('franchise-goalie-records?cayenneExp=franchiseId=', as.character(ID))
    get_db_records(temp_string)}
  else if(!is.null(team)){
    team_alt = tail(strsplit(team, " ")[[1]], n = 1)
    temp_ID <- df_franchise_det %>% filter(toupper(short_name) == toupper(team_alt)) %>% select(id)
    temp_string <- paste0('franchise-goalie-records?cayenneExp=franchiseId=', as.character(temp_ID))
    get_db_records(temp_string)
  }
}

# This pulls skater information
get_skater_data <- function(team = NULL, ID = NULL){
  df_franchise_det <- get_db_records('franchise-detail')
  df_franchise_det <- df_franchise_det$data %>% rowwise() %>%
    mutate(short_name = tail(strsplit(teamFullName, " ")[[1]], n = 1))
  if(is.null(team) & !is.null(ID)){
    temp_string <- paste0('franchise-skater-records?cayenneExp=franchiseId=', as.character(ID))
    get_db_records(temp_string)}
  else if(!is.null(team)){
    team_alt = tail(strsplit(team, " ")[[1]], n = 1)
    temp_ID <- df_franchise_det %>% filter(toupper(short_name) == toupper(team_alt)) %>% select(id)
    temp_string <- paste0('franchise-skater-records?cayenneExp=franchiseId=', as.character(temp_ID))
    get_db_records(temp_string)
  }
}

get_team_stats2 <- function(team = NULL){
  df_franchise_det <- get_db_records('franchise-detail')
  df_franchise_det <- df_franchise_det$data %>% rowwise() %>%
    mutate(short_name = tail(strsplit(teamFullName, " ")[[1]], n = 1))
  
  if(is.numeric(team)){
    temp_string <- paste0('teams/', as.character(team))
    return(get_db_stats(temp_string, '?expand=team.stats'))}
  
  else if(is.character(team)){
    team_alt = tail(strsplit(team, " ")[[1]], n = 1)
    temp_ID <- df_franchise_det %>% filter(toupper(short_name) == toupper(team_alt)) %>% select(id)
    temp_string <- paste0('teams/', as.character(temp_ID))
    temp_df <- get_db_stats(temp_string, '?expand=team.stats')
    temp_df <- temp_df$teams$teamStats[[1]]
    temp_df <- temp_df$splits[[1]]
    return(temp_df)}
  
  else {
    return(get_db_stats('teams'))
  }
}
```

## Exploratory Data Analysis

The functions developed in the prior section will be used to connect to and to pull data from the NHL API service. First, the some datasets will be created using the following commands.

```{r explore, echo = FALSE}
# Exploratory analysis of the data sets
#########################################################################################
#(Returns id, firstSeasonId and lastSeasonId and name of every team in the history of the NHL)
df_franchise <- get_db_records('franchise') 

#(Returns Total stats for every franchise (ex roadTies, roadWins, etc))
df_fran_team_tot <- get_db_records('franchise-team-totals') 

#(Drill-down into season records for a specific franchise)
df_records <- get_db_records('franchise-season-records?cayenneExp=franchiseId=16') 

#(Goalie records for the specified franchise)
df_goalie_bruins <- get_db_records('franchise-goalie-records?cayenneExp=franchiseId=6')

#(Skater records, same interaction as goalie endpoint)
df_skater_bruins <- get_db_records('franchise-skater-records?cayenneExp=franchiseId=6')

# (Franchise details, websites, etc)
df_franchise_det <- get_db_records('franchise-detail')

# Team stats
df_teams <- get_team_stats2()

# Bruin stats
df_bruins <- get_team_stats2('bruins')
```

Here are the column names for all the datasets. In the case of the data sets for a particular team, one was chosen at random.

```{r table names}
# Column names of each dataset
names(df_franchise$data)
names(df_fran_team_tot$data)
names(df_franchise_det$data)
names(df_records$data) #Data for the Flyers
names(df_goalie_bruins$data) #goalie data for the Bruins :(
names(df_skater_bruins$data) #skater data for the Bruins
names(df_teams$teams)
names(df_bruins)
```

Here is some dataframe information

```{r dataframe_info}
# Information for each dataset
str(df_franchise$data)
str(df_fran_team_tot$data)
str(df_franchise_det$data)
str(df_records$data) #Data for the Flyers
str(df_goalie_bruins$data) #goalie data for the Bruins :(
str(df_skater_bruins$data) #skater data for the Bruins
str(df_teams$teams)
str(df_bruins)
```

Summary Statistics of some of the datasets

```{r summary}
# Summaries of each dataset
summary(df_franchise$data)
summary(df_fran_team_tot$data)
summary(df_franchise_det$data)
summary(df_records$data) #Data for the Flyers
summary(df_goalie_bruins$data) #goalie data for the Bruins :(
summary(df_skater_bruins$data) #skater data for the Bruins
summary(df_teams$teams)
summary(df_bruins)
```

## Team Specific Information

Here is a few test scenarios for a handful of team specific datasets. 

```{r}
# Quick Test
df_temp_goalie <- get_goalie_data("Maple Leafs")
df_temp_skater <- get_skater_data("Maple Leafs")
df_temp <- get_records("Maple Leafs")
df_temp$data %>% select(id, franchiseName, lossStreak, winStreak, mostGoals, fewestGoals) %>% knitr::kable()

#names(df_temp_goalie$data)
df_temp_goalie <- get_goalie_data("Maple Leafs")
df_temp_skater <- get_skater_data("Maple Leafs")

names(df_temp_goalie$data)
names(df_temp_skater$data)
```

Creating some stats and frequency tables.

```{r}
df_temp_goalie$data %>% select(firstName, lastName, mostSavesOneGame, gamesPlayed) %>% 
  arrange(desc(gamesPlayed)) %>% head() %>% knitr::kable()

df_temp_skater$data %>% select(positionCode, mostGoalsOneGame) %>% table() %>% knitr::kable()
```

## Setup ID Data

There seems to be descrepency between the number of active teams. This was found to be due to the recent addition of the Seattle Kraken as an expansion team. They are an active team but have not really participated in too many games. 

```{r dataframes, echo = FALSE, message = FALSE, error= FALSE}
df_franchise$data %>% filter(is.na(lastSeasonId)) %>% select(id, teamCommonName) %>% knitr::kable()

a <- df_franchise$data %>% filter(is.na(lastSeasonId)) %>% select(id, fullName)

b <- df_fran_team_tot$data %>% filter(is.na(lastSeasonId) & gameTypeId == 2) %>% select(teamName)

c <- a$fullName[!(a$fullName %in% b$fullName)]

print(c)
```

## Display Data

The following is an example of how to parse and join different data sets and to visualize the data. From the charts, it can be seen that conferences are not based on divisions. Also, that the eastern conference seemed to have formed sooner than the western.

```{r datatables, message = FALSE, error = FALSE}
df_teams$teams %>% select(id, name) %>% nrow()
df_teams$teams %>% filter(active == TRUE) %>% select(id, name) %>% nrow()

# Bar charts based on conference and divisions.
df_teams$teams %>% filter(active == TRUE) %>% 
  ggplot(. ,aes(conference.name, fill = as.factor(division.id))) + geom_bar() + 
  labs(x = "Conference ID", y = "Count") + 
  scale_fill_discrete(name = "Divisions", 
                      labels = c("MassMutual East", 
                                 "Discover Central",
                                 "Honda West",
                                 "Scotia North"))

df_teams$teams %>% filter(active == TRUE) %>% 
  ggplot(. ,aes(firstYearOfPlay)) + geom_bar(aes(fill = conference.name)) + 
  labs(x = "First Year of Play", y = "Count") +
  scale_fill_discrete(name = "Conference")
```

From boxplot data Scotia North division has the highest average number of wins in both conferences. Scotia North has a wider distribution in the Eastern Conference than it does in the Western. Overall, Mass Mutual East has the highest average wins regardless of conference. This information needs further refinement because, number of games played from older divisions/conferences can be skewing the data.

```{r}
# Combining some endpoint data
df_fil_fran_tot <- df_fran_team_tot$data %>% filter(is.na(lastSeasonId) & gameTypeId == 2) %>% rename(abbreviation = triCode)
df_fil_team <- df_teams$teams %>% filter(active == TRUE)

#Rename a column in df_fil_fran_tot to match that of a column in df_fil_team and then inner join them on the newly renamed column

df_com <- df_fil_team %>% inner_join(., df_fil_fran_tot, by = 'abbreviation')

df_com %>% ggplot(., aes(x = division.name, y = wins)) + geom_boxplot(fill = "gray") +
  labs(title = "Distribution of Wins Betwen the Conferences", 
       x = "Conference Name", y = "Number of Wins") + facet_grid(cols = vars(conference.name)) +
  theme(axis.text.x = element_text(angle = 25, hjust=1))

df_com %>% ggplot(., aes(x = division.name, y = wins)) + geom_boxplot(fill = 'gray') + 
  labs(x = "Divisions", y = "Number of Wins")
```

Scatter plots of wins do not seem to point to any relationships in the data. We can see that the Knights have the highest Percentage Points and the Coyotes the lowest. Percent Home Wins versus Total Wins do not have a relationship according to first glance at the scatter plot.

```{r}
df_com %>% ggplot(., aes(x = abbreviation, y = pointPctg)) + geom_point() +
  labs(title = "Percentage Points Versus NHL Team", x = "NHL Teams (abbr)", y = "Point Pctg") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


df_com %>% mutate(win_per_tot = wins / gamesPlayed, win_per_home = homeWins / wins) %>% 
  ggplot(., aes(x = win_per_tot, y = win_per_home)) + geom_point(aes(color = as.factor(division.id))) + 
  labs(x = "Percent Wins from Total Games", y = "Percent Wins Home Games") + 
  scale_color_discrete(name = "Divisions", 
                      labels = c("MassMutual East", 
                                 "Discover Central",
                                 "Honda West",
                                 "Scotia North")) 
```
Displaying Home wins in a Histogram shows heavy right skew but this could be due to teams with longer tentures having played more games. When creating a histogram base on Home Wins as a percentage of total wins, a more even distribution is observed.

```{r}
df_fran_team_tot$data %>% ggplot(., aes(homeWins)) + geom_histogram(aes(y = ..density..), bins = 15) + 
  geom_density(adjust = 0.75, color = 'red', size = 0.25, outline.type = 'full') + 
  labs(x = "Number of Home Wins", title = "Histogram of Team Home Wins")

df_com %>% mutate(win_per_home = homeWins / wins) %>% 
  ggplot(., aes(win_per_home)) + geom_histogram(aes(y = ..density..), bins = 15) + 
  geom_density(adjust = 0.75, color = 'red', size = 0.25, outline.type = 'full') +
  labs(x = "Home Wins Percentage", title = "Histogram of Home Wins as a Percentage of Total Wins")
```

## One Stop Shop Data

If one does not want to run individual functions the grab_all function can be run instead. The function uses 5 arguments:

- all: to return all nhl data 
- team: to return specific team data 
- ID: same as team but uses the ID number instead 
- goalie: to return goalie information 
- skater: to return skater information. 
- The output for the function is a named list containing all the desired information.  

<Font color = red>**WARNING** Named list can be large! </font>.

```{r one_stop, message = FALSE, error= FALSE}

grab_all <- function(all = FALSE, team = NULL, ID = NULL, goalie = FALSE, skater = FALSE){
  if(all == TRUE){
    df_franchise <- get_db_records('franchise') 
    df_fran_team_tot <- get_db_records('franchise-team-totals')
  } else if(all == FALSE) {
    df_franchise <- NULL
    df_fran_team_tot <- NULL}
 
  if(!is.null(team) & is.null(ID)){
    df_team_data <- get_records(team)
    df_team_stats <- get_team_stats2(team)
  } else if(is.null(team) & !is.null(ID)) {
    df_team_data <- get_records(ID = ID)
    df_team_stats <- get_team_stats2(ID)
  } else if(!is.null(team) & !is.null(ID)) {
    df_team_data <- get_records(team)
    df_team_stats <- get_team_stats2(team)
  } else if(is.null(team) & is.null(ID)) {
    df_team_data <- NULL
    df_team_stats <- NULL
  }
  
  if(goalie == TRUE & !is.null(team) & is.null(ID)){
    df_temp_goalie <- get_goalie_data(team)
  } else if(goalie == TRUE & is.null(team) & !is.null(ID)){
    df_temp_goalie <- get_goalie_data(ID = ID)
  } else if(goalie == TRUE & is.null(team) & is.null(ID)){
    df_temp_goalie <- get_goalie_data(team)
  } else if(goalie == FALSE){
    df_temp_goalie <- NULL
  }
  
  if(skater == TRUE & !is.null(team) & is.null(ID)){
    df_temp_skater <- get_skater_data(team)
  } else if(skater == TRUE & is.null(team) & !is.null(ID)){
    df_temp_skater <- get_skater_data(ID = ID)
  } else if(skater == TRUE & is.null(team) & is.null(ID)){
    df_temp_skater <- get_skater_data(team)
  } else if(skater == FALSE){
    df_temp_skater <- NULL
  }
  
  df_list <- list("franchises" = df_franchise, "team total" = df_fran_team_tot, 
                  "goalie information" = df_temp_goalie, 
                  "skater information" = df_temp_skater, 
                  "team data" = df_team_data, 
                  "team stats" = df_team_stats)
  
  df_list <- df_list[-which(sapply(df_list, is.null))]
  
  return(df_list)
}

# Test grab_all function
grab_all(all = TRUE, team = 'stars')
```